### 2016-09-23 proposal.tex
init

\documentclass[11pt,letterpaper]{article}
\usepackage[margin=1in]{geometry}

\renewcommand{\baselinestretch}{1.2}

\pagenumbering{gobble}

\title{{\Large Integrated Least Effort Knowledge Networks}}
\author{Lucas E. Morales \texttt{\{lucasem\}}}
\date{}

\begin{document}

\maketitle

\section{Introduction}

Knowledge, as a familiarity with things, is attained by experience via
learning. Learning mechanisms both construct and consume knowledge, allowing
for higher order learning compared to the base of sensory input. This
interaction must not be immediately free to use all knowledge: a particular
context must motivate certain atoms of knowledge to be readily available
while setting others further apart, depending on their relationships to
those in contextual proximity.

Knowledge can be represented as distinct atoms of information and their
relations in the structure of a connected network. The network is
constructed by preferential ``least effort'' attachment \cite{cancho03},
where the learning of a new atom of knowledge joins the network with
relations to the contextual knowledge from which it was learned. This
network conforms to a mathematical pattern known as Zipf's Law
\cite{zipf49}, where the degree of connectivity for nodes in the network
follow a class of power law distributions \cite{barabasi99}. A similar
pattern arises in social networks, the webgraph of the Internet, and the
statistics of linguistic distribution \cite{clauset09}.

% Creation of such a system, built on existing learning mechanisms, requires
% the completion of subtasks as follows:
% 
% \begin{itemize}
%   \setlength\itemsep{0em}
%   \item Design the knowledge network to permit a useful interface for a set
%     of sufficient learning mechanisms.
%   \item Augment existing learning mechanisms to utilize contextual
%     knowledge.
%   \item Augment existing learning mechanisms to create and update
%     information in the knowledge network.
%   \item Connect distinct kinds of knowledge generated by multiple learning
%     mechanisms in some manner, at least including temporal locality, and
%     perhaps including meta-interpretive similarities.
% \end{itemize}

\section{Relevant Works}

In the discipline of artificial intelligence, much effort is placed on
designing accurate learning mechanisms.
These learning mechanisms are generally either designed for a particular
domain, or
designed to specialize given consistent domain-specific input.

Special-purpose learning mechanisms include visual object recognition
systems based on the human visual cortex \cite{serre07}, program learning
\cite{liang10}, visual feature modification of images \cite{kulkarni15}, and
sound texture perception and synthesis \cite{mcdermott11}.

Specializable learning mechanisms include hierarchical Bayesian methods
\cite{tenenbaum01}, inductive logic programming \cite{lavrac94}, generative
adversarial networks \cite{goodfellow14}, and search on a context-free
grammar \cite{piantadosi12}.

Each of these mechanisms are remarkable in a specified manner, but lack the
generality that is apparent in human cognition. This is the problem of
\emph{knowledge}, and most learning mechanisms have some internal
representation for it. Those representations are mechanism-specific, and
generally have a flat hierarchy (i.e. where all knowledge is, in some
manner, equidistant). A flat representation of knowledge is flawed in
modeling cognition because it reduces capability of specialization in
multiplicity by making all knowledge focused in the single domain that the
learning mechanism is confined in. Attempts have not been made in alignment
with the goals of the knowledge network this proposal aims to build, to
design and augment learning mechanisms around \emph{contextual} knowledge in
a manner that is inter-operable between distinct learning mechanisms.

\begin{thebibliography}{}
  \bibitem{cancho03} i Cancho, R. F., \& Ricard V. Solé. (2003)
    Least effort and the origins of scaling in human language.
    In \emph{Proceedings of the National Academy of Sciences} 100(3), 788-791.
  \bibitem{zipf49} Zipf, G. K. (1949).
    Human Behaviour and the Principle of Least-Effort.
  \bibitem{barabasi99}
    Barab\'asi, A. L., \& Albert, R. (1999).
    Emergence of scaling in random networks.
    \emph{Science}, 286(5439), 509-512.
  \bibitem{clauset09} Clauset A., Shalizi, C. R., \& Newman M. E. J.  (2009)
    Power-law distributions in empirical data.
    \emph{SIAM Review} 51(4), 661-703 (arXiv:0706.1062, doi:10.1137/070710111)
%  \bibitem{lake15} Lake, B. M., Salakhutdinov, R., \& Tenenbaum, J. B. (2015).
%    Human-level concept learning through probabilistic program induction.
%    \emph{Science}, 350(6266), 1332-1338.
  \bibitem{serre07}
    Serre, T., Oliva, A., \& Poggio, T. (2007).
    A feedforward architecture accounts for rapid categorization.
    In \emph{Proceedings of the National Academy of Sciences}, 104(15), 6424-6429.
  \bibitem{liang10}
    Liang, P., Jordan, M. I., \& Klein, D. (2010).
    Learning programs: a hierarchical Bayesian approach.
    \emph{International Conference on Machine Learning}: 639–646
  \bibitem{kulkarni15}
    Kulkarni, T. D., Whitney, W. F., Kohli, P., \& Tenenbaum, J. (2015).
      Deep convolutional inverse graphics network.
      In \emph{Advances in Neural Information Processing Systems} (pp. 2539-2547).
  \bibitem{mcdermott11}
    McDermott, J. H., \& Simoncelli, E. P. (2011).
      Sound texture perception via statistics of the auditory periphery: evidence from sound synthesis.
      \emph{Neuron}, 71(5), 926-940.
  \bibitem{tenenbaum01}
    Tenenbaum, J. B., \& Griffiths, T. L. (2001).
      The rational basis of representativeness.
      In \emph{Proceedings of the 23rd annual conference of the Cognitive
      Science Society} (p. 103641).
  \bibitem{lavrac94}
    Lavrac, N., \& Dzeroski, S. (1994).
      Inductive Logic Programming.
      In \emph{WLP} (pp. 146-160).
  \bibitem{goodfellow14}
    Goodfellow, I., Pouget-Abadie, J., Mirza, M., Xu, B., Warde-Farley, D.,
      Ozair, S., ... \& Bengio, Y. (2014).
      Generative adversarial nets.
      In \emph{Advances in Neural Information Processing Systems} (pp. 2672-2680).
  \bibitem{piantadosi12}
    Piantadosi, S. T., Tenenbaum, J. B., \& Goodman, N. D. (2012).
      Bootstrapping in a language of thought: A formal model of numerical concept learning.
      \emph{Cognition}, 123(2), 199-217.
\end{thebibliography}

\end{document}